---
title: "STA 613 Report Markdown Draft"
author: "Shijia Bian"
date: "May 3, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load the cleaned up date set `dummy.freq.Table.Rdata`, and write down the marks for the breed. 

dummy.freq.Table.Rdata has the column be the allele frequency for each of the species, the breed mark is the breed specification for each of the species.

```{r}
load("~/Dropbox/Duke Statistics Courses/Spring 2016 Duke Course/STA 613 COMSCI BIO/Final Project/Data/dummy.freq.Table.Rdata")

breed = c("wildCanids", "wildCanids", 
          "wildCanids", "wildCanids",
          "wildCanids", "wildCanids",
          "acientBreeds", "acientBreeds",
          "acientBreeds", "acientBreeds",
          "acientBreeds", "acientBreeds",
          "acientBreeds", "acientBreeds", 
          "acientBreeds", "acientBreeds",
          "acientBreeds", "acientBreeds",
          "spitz", "spitz",
          "spitz", "spitz",
          "spitz", "spitz",
          "spitz", "spitz",
          "spitz", "spitz",
          "spitz", "spitz",
          "spitz", "spitz",
          "toy", "toy", 
          "toy", "toy", 
          "herding", "herding",
          "herding", "herding",
          "herding", "herding",
          "herding", "herding",
          "herding", "herding",
          "mastiff", "mastiff",
          "mastiff", "mastiff",
          "mastiff", "mastiff",
          "mastiff", "mastiff",
          "mastiff", "mastiff",
          "mastiff", "mastiff",
          "mastiff", "mastiff",
          "mastiff", "mastiff",
          "mastiff", "mastiff",
          "mastiff", "mastiff",
          "mastiff", "mastiff",
          "mastiff", "mastiff",
          "mastiff", "mastiff",
          "mastiff", "mastiff",
          "mastiff", "mastiff")


breadMark = c()
for (i in 1:38) {
  breadMark.du = breed[2*i]
  breadMark = c(breadMark, breadMark.du)
}
```


calculate the population mean for each of the population, and the overall mean
```{r}
wildCanidsMean = apply(dummy.freq.Table[,breadMark=="wildCanids"], 1, mean)
acientBreedsMean = apply(dummy.freq.Table[,breadMark=="acientBreeds"], 1, mean)
spitzMean = apply(dummy.freq.Table[,breadMark=="spitz"], 1, mean)
toyMean = apply(dummy.freq.Table[,breadMark=="toy"], 1, mean)
herdingMean = apply(dummy.freq.Table[,breadMark=="herding"], 1, mean)
mastiffMean = apply(dummy.freq.Table[,breadMark=="mastiff"], 1, mean)
overallMean = apply(dummy.freq.Table, 1, mean)
```

Calculate the covariance matrix 
```{r}
popMean = cbind(wildCanidsMean, acientBreedsMean, spitzMean, toyMean, herdingMean, mastiffMean)
dim(popMean) # 38312     6
popMean.matrix = as.matrix(popMean)
cov(popMean.matrix)
```

```{r}
library("mmod")
data("nancycats")
nancycats

#install.packages("poppr")
library("poppr")

#install.packages("ape")
library("ape") # To visualize the tree using the "nj" function

#install.packages("magrittr")
library("magrittr")
data(microbov)
set.seed(10)
ten_samples <- sample(nInd(microbov), 10)
mic10       <- microbov[ten_samples]
(micdist    <- provesti.dist(mic10))
```

```{r}
allelFeq = as.matrix(dummy.freq.Table)
allelFeq.t = t(allelFeq)
allelFeq.tt=na.omit(allelFeq.t)

allelFeq.ttP = allelFeq.tt
clusters <- hclust(dist(allelFeq.ttP), method = 'average')
plot(clusters)

```

composite liklihood
```{r}
# https://cran.r-project.org/web/packages/CompRandFld/CompRandFld.pdf
# page 11
#install.packages("CompRandFld")
library(CompRandFld)

#install.packages("RandomFields")
library(RandomFields)

# install.packages("scatterplot3d")
library(scatterplot3d)
set.seed(31231)
# Set the coordinates of the points:
x <- runif(100, 0, 10)
y <- runif(100, 0, 10)
coords<-cbind(popMean[,1],popMean[,2])[1:1000,]
################################################################
###
### Example 1. Plot of covariance and variogram functions
### estimated from a Gaussian random field with exponent
### correlation. One spatial replication is simulated.
###
###
###############################################################
# Set the model's parameters:
corrmodel <- "exponential"
mean <- 0
sill <- 1
nugget <- 0
scale <- 2
# Simulation of the Gaussian random field:
data <- RFsim(coordx=coords, corrmodel=corrmodel, param=list(mean=mean,
sill=sill, nugget=nugget, scale=scale))$data
# Maximum composite-likelihood fitting of the Gaussian random field:

start<-list(scale=scale,sill=sill,mean=mean(data))
fixed<-list(nugget=nugget)
# Maximum composite-likelihood fitting of the random field:
fit <- FitComposite(data, coordx=coords, corrmodel=corrmodel,likelihood="Marginal",
type="Pairwise",start=start,fixed=fixed,maxdist=6)
# Results:
print(fit)
# Empirical estimation of the variogram:
vario <- EVariogram(data, popMean[,1][1:1000], popMean[1:1000])
# Plot of covariance and variogram functions:
par(mfrow=c(1,2))
Covariogram(fit, show.cov=TRUE, show.range=TRUE,
show.vario=TRUE, vario=vario,pch=20)


################################################################
##
### Example 2. Plot of covariance and extremal coefficient
### functions estimated from a max-stable random field with
### exponential correlation. n idd spatial replications are
### simulated.
###
###############################################################
set.seed(1156)
# Simulation of the max-stable random field:
data <- RFsim(coordx=coords, corrmodel=corrmodel, model="ExtGauss", replicates=20,
param=list(mean=mean,sill=sill,nugget=nugget,scale=scale))$data
start=list(sill=sill,scale=scale)
# Maximum composite-likelihood fitting of the max-stable random field:
fit <- FitComposite(data, coordx=coords, corrmodel=corrmodel, model='ExtGauss',
replicates=20, varest=TRUE, vartype='Sampling',
margins="Frechet",start=start)
```


The Block Design

Equation 20, estimate G

```{r}
dim(dummy.freq.Table)
#### divide the table into 8 blocks
popMean = cbind(wildCanidsMean, acientBreedsMean, spitzMean, toyMean, herdingMean, mastiffMean)
dim(popMean) # 38312     6
popMean.matrix = as.matrix(popMean)
cov(popMean.matrix)
```

Equation 24, estimate W hat:

```{r}
w_ij = cov(popMean.matrix)
col_mean = apply(popMean.matrix, 1, mean)
G_ij = t(popMean.matrix - col_mean)  %*% (popMean.matrix - col_mean)
```

Equation 27, estimate sigma ij hat:

Divide the popMean.matrix into 4 blocks
```{r}
dim(popMean.matrix)
seq(1, 38312, length.out = 5)

popMean.matrix1 = popMean.matrix[1:9578, ]
popMean.matrix2 = popMean.matrix[9579:19156, ]
popMean.matrix3 = popMean.matrix[19157:28734, ]
popMean.matrix4 = popMean.matrix[28735:38312, ]

block1 = cov(popMean.matrix1)
block2 = cov(popMean.matrix2)
block3 = cov(popMean.matrix3)
block4 = cov(popMean.matrix4)

w_hat_ij = (block1+block2+block3+block4)/4

sigma_ij_square = ((block1-w_hat)^2+(block2-w_hat)^2+(block3-w_hat)^2+(block4-w_hat)^2)/(4*3)
```

Eq 28
```{r}
overall_Exp = mean(w_hat_ij)
w_ij = matrix(overall_Exp, 6, 6)
temp = 1

for (i in 1:dim(w_ij)[1]) {
  for (j in 1:dim(w_ij)[2]) {
    stor = pnorm(w_hat_ij[i,j], w_ij[i,j], sqrt(sigma_ij_square[i,j]))
    if (stor != 0) {
      temp = temp * stor
      print(temp)
    }  
  }
}
temp
```

